---
title: "scRNAseq_workflow_protocol (supplementary mouse)"
author: "Wenyan Li"
date: '2023-11-17'
output:

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up working directory
```{r}
# load required packages
library(rstudioapi)
library(Seurat)
library(patchwork)
library(S4Vectors)
library(tidyverse)
library(ggplot2)

# set a seed to re-produce pseudorandom numbers
set.seed(99)

# Obtain the path of this rmarkdown file and assign it to object "get_path"
get_path <- dirname(rstudioapi::getSourceEditorContext()$path)

# Set "get_path" as Working Directory
setwd(get_path)

# Save images and data to this folder
saveTo <- "Supplementary mouse output"
```

# Pre-processing

## create a function to save images 
```{r}
# Create a function to save output images

func_save_images <- function(image.object, # plot objects
                             image.name, # name of the plot
                             image.path, # path to save image
                             h = 8, # image height in inch
                             w = 15, # image width in inch
                             r = 300, # image resolution
                             isHeatmap = F # check if image is a Seurat heatmap
){
  
  ifelse(!dir.exists(image.path), 
         dir.create(image.path,
                    recursive = T), 
         FALSE)
  
  if(isHeatmap){
    
      for(i in seq_along(image.name)){
        temp.path <- paste(image.path,
                           image.name[i], 
                           sep = "/")
        
        tiff(filename = paste(temp.path,
                              "tiff",
                              sep = "."),
             width = w, 
             height = h,
             res = r, 
             units = "in")
        
        image.object[[i]]
        dev.off()
      } # end of for 1st loop
    
  }else{
    
      for(i in seq_along(image.name)){
        temp.path <- paste(image.path,
                           image.name[i], 
                           sep = "/")
        
        ggsave(filename = paste(temp.path,
                                "tiff",
                                sep = "."), 
               plot =  image.object[[i]],
               width = w, 
               height = h, 
               units = "in", 
               dpi = r, 
               limitsize = FALSE)
        
      } # end of for 2nd loop
    
  } # end of if isHeatmap statement
  
}# end of func_save_images function

```


## Load Seurat object
```{r}
# Choose a folder/experiment to analyse
exp1 <- "BD-Demo-mouse-wta"

# Load the Seurat objects as demo_seurat_1 and demo_seurat_2
demo_seurat_1 <- readRDS(list.files(exp1, full.names = T, pattern = "*.rds"))

```


## Check Seurat object
```{r}
demo_seurat_1
```

## Data normalisation, scaling and finding clusters
```{r}
# Build function

# The size of AbSeq panels varies, reduce the number for ab_pc_num if your panel is smaller than 10.
func_quick_process <- function(demo_seurat,
                               ab_pc_num = 10, # number of PCA components to use for protein
                               rna_pc_num = 15, # number of PCA components to use for RNA
                               ab_reduction_res = 0.8, # cluster resolution for protein
                               rna_reduction_res = 0.8) # cluster resolution for RNA
  
     {
  
      # check if the Seurat object has protein assay
      if("AB" %in% Seurat::Assays(demo_seurat))
        
          {  
           Seurat::DefaultAssay(demo_seurat) <- 'AB'
  
           # Normalize and scale data
           demo_seurat <- demo_seurat %>% 
                          Seurat::NormalizeData(normalization.method = 'CLR', 
                                                margin = 2)  %>% 
                          Seurat::FindVariableFeatures() %>% 
                          Seurat::ScaleData() 

	         # perform PCA
           demo_seurat <- Seurat::RunPCA(object = demo_seurat, 
                                         reduction.name = 'apca')

	         # perform UMAP
           demo_seurat <- Seurat::RunUMAP(demo_seurat, 
                                          reduction = 'apca', 
                                          dims = 1:ab_pc_num, 
                                          assay = 'AB', 
                                          reduction.name = 'adt.umap',
			                                    reduction.key = 'adtUMAP_')

	         # Find clusters
           demo_seurat <- Seurat::FindNeighbors(demo_seurat, 
                                                reduction = "apca", 
                                                dims = 1:ab_pc_num)

           demo_seurat <- Seurat::FindClusters(demo_seurat, 
                                               resolution = ab_reduction_res, 
                                               graph.name = "AB_snn")

           } # end of if statement
  
     # Change default assay to RNA
     Seurat::DefaultAssay(demo_seurat) <- "RNA"

     # Calculate percentages of mitochondrial gene expression for every cell. If it is a
     #  targeted sequencing, then the output of this commend line is all zeros.
     demo_seurat <- Seurat::PercentageFeatureSet(demo_seurat, 
                                                 pattern = "^mt[-|.]", 
                                                 col.name = "percent.mt") 

     # find top most variant genes
     demo_seurat <- demo_seurat %>% 
                    Seurat::NormalizeData() %>% 
		                Seurat::FindVariableFeatures(., 
		                                             selection.method = "vst")

     # scale data
     demo_seurat <- Seurat::ScaleData(demo_seurat, 
                                      verbose = FALSE)

     # perform PCA
     demo_seurat <- Seurat::RunPCA(demo_seurat, 
                                   npcs = rna_pc_num, 
                                   verbose = FALSE)

     # perform UMAP
     demo_seurat <- Seurat::RunUMAP(demo_seurat, 
                                    reduction = "pca", 
                                    dims = 1:rna_pc_num)

     # Find clusters
     demo_seurat <- Seurat::FindNeighbors(demo_seurat, 
                                          reduction = "pca", 
                                          dims = 1:rna_pc_num)

     demo_seurat <- Seurat::FindClusters(demo_seurat, 
                                         resolution = rna_reduction_res)

     demo_seurat <- Seurat::BuildClusterTree(demo_seurat)

     # return Seurat object as output
     return(demo_seurat)

} # end of func_quick_process function

# Use the above function to process the Seurat objects
demo_seurat_1 <- func_quick_process(demo_seurat_1)

```


# Save raw Seurat object
```{r}
# Create a folder called saveTo
ifelse(!dir.exists(saveTo), 
       dir.create(saveTo, 
                  recursive = T), 
       FALSE)

# Save Seurat objects to local drive
save(demo_seurat_1,
     file = paste(saveTo, 
                  "raw demo Seurat objects.RData", 
                  sep = "/"))

```


## Results: Data normalization, scaling and finding clusters
```{r}
# QC plots – check mitochondrial gene percentages
p1 <- Seurat::VlnPlot(demo_seurat_1, features = "percent.mt", 
              group.by = "seurat_clusters") + 
              Seurat::NoLegend() + 
              ggtitle("mouse demo exp 1 - MT gene%")

func_save_images(image.object = list(p1),
                 image.name = "Mouse Demo 1 - Initial MT genes percentages",
                 image.path = saveTo,
                 h = 5, 
                 w = 5, 
                 r = 300, 
                 isHeatmap = F
                 )

p2 <- Seurat::FeatureScatter(demo_seurat_1, 
                             feature1 = "nCount_RNA", 
                             feature2 = "nFeature_RNA", 
                             group.by = "seurat_clusters") + 
              scale_x_log10() +
              scale_y_log10() +
              ggtitle("mouse demo exp 1")

func_save_images(image.object = list(p2),
                 image.name = "Initial nCounts and nFeatures",
                 image.path = saveTo,
                 h = 5, 
                 w = 5, 
                 r = 300, 
                 isHeatmap = F
                 )

p1 
p2

```


## Remove poor quality cells
```{r}
# filter out cells with MT genes percentage > 50 (%) and 
# cells with low nFeature_RNA
subset_demo_seurat_1 <- subset(demo_seurat_1, 
                               subset = percent.mt < 30 & 
                                        nFeature_RNA > 200, 
                               invert = F)

p3 <- Seurat::VlnPlot(subset_demo_seurat_1, 
                      features = "percent.mt", 
                      group.by = "seurat_clusters") + 
              Seurat::NoLegend() + 
              ggtitle("mouse demo exp 1 - MT gene%")

func_save_images(image.object = list(p3),
                 image.name = "Subset - MT genes percentages (before re-clustering)",
                 image.path = saveTo,
                 h = 5, 
                 w = 5, 
                 r = 300, 
                 isHeatmap = F
                 )

p4 <- Seurat::FeatureScatter(subset_demo_seurat_1, 
                             feature1 = "nCount_RNA", 
                             feature2 = "nFeature_RNA", 
                             group.by = "seurat_clusters") + 
              scale_x_log10() +
              scale_y_log10() +
              ggtitle("mouse demo exp 1")

func_save_images(image.object = list(p4),
                 image.name = "Subset - nCounts and nFeatures (before re-clustering)",
                 image.path = saveTo,
                 h = 5, 
                 w = 5, 
                 r = 300, 
                 isHeatmap = F
                 )

p3 
p4

```


## Re-cluster subsetted samples
```{r}
subset_demo_seurat_1 <- func_quick_process(subset_demo_seurat_1)

```


```{r}
# UMAP plots
p5 <- Seurat::DimPlot(subset_demo_seurat_1, 
                      reduction = "umap", 
                      group.by = "seurat_clusters") + 
              ggtitle("mouse demo exp 1")

func_save_images(image.object = list(p5),
                 image.name = "Subset - UMAP group by clusters",
                 image.path = saveTo,
                 h = 5, 
                 w = 5, 
                 r = 300, 
                 isHeatmap = F
                 )

p5

```


## Save subsetted Seurat objects
```{r}
# Save Seurat objects to local drive
save(subset_demo_seurat_1,
     file = paste(saveTo, 
                  "subsetted demo Seurat objects.RData", 
                  sep = "/"))

```


# Advanced analysis

## load packages
```{r}

# load required packages
library(rstudioapi)
library(Seurat)
library(ggplot2)
library(tidyverse)
library(patchwork)
library(celldex)
library(SingleR)
library(harmony)
library(DoubletFinder)
library(SeuratWrappers)
library(monocle3)
library(slingshot)
library(colorRamps)
library(CellChat)

```


## cell type annotation
```{r}

# cell type annotation
mm_ref <- celldex::ImmGenData()

# build function
func_get_annotation<- function(input_seurat)
  
     {
      DefaultAssay(input_seurat) <- "RNA"
      
      expr_matrix <- GetAssayData(input_seurat, 
                                  slot = "data", 
                                  assay = "RNA")
      
      cluster_id <- input_seurat@meta.data$seurat_clusters
  
      # Optional: annotate cells by groups
      prediction_by_cluster <-SingleR::SingleR(test = expr_matrix,
                                       ref = mm_ref, # make sure reference data is correct
                                       labels = mm_ref$label.main, # make sure reference data is correct
                                       clusters = cluster_id)

      # Annotation cells individually  
      prediction_by_cell <- SingleR::SingleR(test = expr_matrix,
                                     ref = mm_ref,  # make sure reference data is correct
                                     labels = mm_ref$label.main) # make sure reference data is correct
      
      # Save SingleR results to the Seurat object
      input_seurat@misc[["SingleR_results"]] <- prediction_by_cell

      # Annotation results
      cell_labels <- prediction_by_cell$labels

      names(cell_labels) <- rownames(prediction_by_cell$labels)

      # add annotation information to Seurat object under meta.data
      input_seurat <- AddMetaData(input_seurat, 
                                  metadata = cell_labels,
                                  col.name = "cell_type")
      
      # make cell types with less than 10 cells as "unknown"
      temp <- table(input_seurat$cell_type)[table(input_seurat$cell_type) < 10] %>% 
              names()
      
      input_seurat$cell_type[input_seurat$cell_type %in% temp] <- "unknown"

      # return Seurat object as output
      return(input_seurat)
      
} # end of func_get_annotation function

# use function to perform singleR cell type annotation
subset_demo_seurat_1 <- func_get_annotation(subset_demo_seurat_1)

```


## Results: Cell type annotation with SingleR
```{r}
# Cell type summary
table(subset_demo_seurat_1$cell_type)

```


```{r}
# Cell annotation scores
p_cell_1 <- plotScoreHeatmap(subset_demo_seurat_1@misc$SingleR_results,
                             show_colnames = F)

func_save_images(image.object = list(p_cell_1),
                 image.name = c("Subset - heatmap of singleR annotation scores - demo 1"),
                 image.path = saveTo,
                 h = 5,
                 w = 6,
                 r = 300,
                 isHeatmap = F
                 )

# Display cells in UMAP plot
p_cell_2 <- Seurat::DimPlot(subset_demo_seurat_1, 
                            group.by = "cell_type") + 
            ggtitle(Project(subset_demo_seurat_1))

func_save_images(image.object = list(p_cell_2),
                 image.name = "Subset - UMAP group by cell type",
                 image.path = saveTo,
                 h = 5, 
                 w = 6, 
                 r = 300, 
                 isHeatmap = F
                 )

p_cell_1 
p_cell_2

```


##Find doublets with DoubletFinder
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Find doublets

# BD provided doublet rates with different cell load numbers
rhapsody_doublet_rate <- data.frame(
                         "cell_num" = c(100,500,1000*(1:20)), 
                         "rate" = c(0, 0.1, 0.2, 0.5, 0.7, 1, 
                                    1.2, 1.4, 1.7, 1.9, 2.1, 
                                    2.4, 2.6, 2.8, 3.1, 3.3, 
                                    3.5, 3.8, 4, 4.2, 4.5 , 4.7))

# Build a linear model to calculate theoretical doublet rate
model_rhap <- lm(rate ~ cell_num, 
                 rhapsody_doublet_rate)

# define function
func_get_doublets <- function(seuratObj,
                              est_doublet_model = model_rhap,
                              pc = 1:15) # number of PC components to be used
     {
  
     DefaultAssay(seuratObj) <- "RNA"
  
     # Find pK values
     sweep.res.list <- paramSweep_v3(seuratObj, 
                                     PCs = pc, 
                                     sct = F)

     sweep.stats <- summarizeSweep(sweep.res.list, 
                                   GT = FALSE)
  
     bcmvn <- find.pK(sweep.stats)

     pK_bcmvn <- bcmvn$pK[which.max(bcmvn$BCmetric)] %>% 
                 as.character() %>% 
                 as.numeric()
  
     # estimate doublet rate based on cell number
     DoubletRate = predict(est_doublet_model, 
                   data.frame(cell_num = dim(seuratObj)[2]))/100
  
     nExp_poi <- round(DoubletRate*ncol(seuratObj)) 
     
     seuratObj <- doubletFinder_v3(seuratObj, 
                                   PCs = pc, 
                                   pN = 0.25, 
                                   pK = pK_bcmvn, 
                                   nExp = nExp_poi, 
                                   reuse.pANN = F, 
                                   sct = F)
  
     temp1 <- grepl("DF.classifications", 
                    colnames(seuratObj@meta.data), 
                    ignore.case = T)

     colnames(seuratObj@meta.data)[temp1] <- "doublet_check"

     seuratObj$doublet_check <- seuratObj$doublet_check
  
     # return output
     return(seuratObj)

} # end of function

# use function to find doublets
subset_demo_seurat_1 <- func_get_doublets(subset_demo_seurat_1,
                                          pc = 1:15)

```


## Results: Find doublets with DoubletFinders
```{r}
# Visualize the result
p_cell_3 <- DimPlot(subset_demo_seurat_1, 
                    group.by = "doublet_check") + 
            ggtitle(Project(subset_demo_seurat_1))

func_save_images(image.object = list(p_cell_3),
                 image.name = "Subset - doublets",
                 image.path = saveTo,
                 h = 5, 
                 w = 6, 
                 r = 300, 
                 isHeatmap = F
                 )

p_cell_3

```



# Finding marker genes
```{r}

# Build function
func_get_marker_genes <- function(input_seurat,
                                  p_adj_cutoff = 0.05,
                                  log2FC_cutoff = 1,
                                  view_top_X_genes = 5)
  
     {
      # Find marker genes for each cluster group against the rest
      Seurat::Idents(input_seurat) <- "seurat_clusters"

      cluster_DGE <- SeuratWrappers::RunPrestoAll(input_seurat, 
                                                  assay = "RNA", 
                                                  only.pos = FALSE, 
                                                  verbose = FALSE)

      # Find marker genes for each cell type against the rest
      Seurat::Idents(input_seurat) <- "cell_type"

      Cell_type_DGE <- SeuratWrappers::RunPrestoAll(input_seurat, 
                                                    assay = "RNA", 
                                                    only.pos = FALSE, 
                                                    verbose = FALSE)

      # example of taking the top X genes in each DGE group and removing the duplicates
      cluster_DGE <- cluster_DGE[abs(cluster_DGE$avg_log2FC) > log2FC_cutoff & 
                             cluster_DGE$p_val_adj < p_adj_cutoff, ]
  
      Cell_type_DGE <- Cell_type_DGE[abs(Cell_type_DGE$avg_log2FC) > log2FC_cutoff & 
                                 Cell_type_DGE$p_val_adj < p_adj_cutoff, ]
      
      top_genes_cluster <- cluster_DGE %>% 
                           group_by(cluster)%>% 
                           slice_max(n = view_top_X_genes, 
                                     order_by = avg_log2FC) %>% 
                           dplyr::pull(gene) %>% 
                           unique()

      top_genes_type <- Cell_type_DGE %>% 
                        group_by(cluster)%>% 
                        slice_max(n = view_top_X_genes, 
                                  order_by = avg_log2FC) %>% 
                        dplyr::pull(gene) %>% 
                        unique()

      # return output as a list
      return(list(DGEs_cluster = cluster_DGE, 
                  DGEs_cell = Cell_type_DGE, 
                  top_cluster_gene = top_genes_cluster, 
                  top_cell_gene = top_genes_type))

} # end of function

# use function to get marker genes
subset_demo1_DGEs <- func_get_marker_genes(subset_demo_seurat_1,
                                           p_adj_cutoff = 0.05,
                                           log2FC_cutoff = 1,
                                           view_top_X_genes = 5) 

```


## Results: Finding marker genes
```{r}
# display example results
subset_demo1_DGEs$top_cell_gene

```


```{r}
# visualise top genes on dotplot
p_cell_4 <- DotPlot(subset_demo_seurat_1, 
                     features = subset_demo1_DGEs$top_cell_gene, 
                     group.by = "cell_type") + 
             coord_flip() +
             RotatedAxis() +
             ggtitle("mouse Subset Demo 1")

func_save_images(image.object = list(p_cell_4),
                 image.name = "Subset - top genes dot plot",
                 image.path = saveTo,
                 h = 12, 
                 w = 7, 
                 r = 300, 
                 isHeatmap = F
                 )

p_cell_4

```



## Trajectory analysis - Monocle3

```{r}
# Monocle3 trajectory analysis

# Define function
func_monocle3 <- function(input_seurat,
                          rna_pc_num = 50)
     {
      # covert Seurat to SingleCellExperiment object
      demo_cds <- SeuratWrappers::as.cell_data_set(input_seurat) 
      
      rowData(demo_cds)$gene_short_name <- rownames(input_seurat)
      
      demo_cds <- preprocess_cds(demo_cds, 
                                 num_dim = rna_pc_num)

      demo_cds <- monocle3::reduce_dimension(demo_cds)
      
      demo_cds <- monocle3::cluster_cells(cds = demo_cds, 
                                          cluster_method = "louvain")

      demo_cds <- monocle3::learn_graph(demo_cds, 
                                        use_partition = TRUE)

      Seurat::Idents(input_seurat) <- "cell_type"

      # This is just an example. Please define root cells that are biologically 
      # correct for your experiment.
      # Note: if ids is an empty object, then this function may generate error.
      # You can use this line instead:       

    	   # demo_cds <- monocle3::order_cells(demo_cds, reduction_method = "UMAP")

      # and remove the following three lines, 
      # "ids <- ..",
      # "root_id <- .." 
      # "demo_cds <- .."

      # for more information, please visit 
      # https://cole-trapnell-lab.github.io/monocle3/docs/trajectories/

      ids <- unique(input_seurat$cell_type) %>% 
             .[grep(paste("CMP", 
                          "GMP", 
                          "HSC", 
                          "CD34", 
                          "stem", 
                          "iPS",
                          sep = "|"),
                    ., 
                    ignore.case = T)]
      
      root_id <- colnames(subset(input_seurat, 
                          idents = ids))

      demo_cds <- monocle3::order_cells(demo_cds, 
                                        reduction_method = "UMAP", 
                                        root_cells = root_id)
      
      colData(demo_cds)$m3_pseudotime <- pseudotime(demo_cds)

      # return Seurat object
      return(demo_cds)
}

# use function to get results
subset_demo_cds_1 <- func_monocle3(subset_demo_seurat_1,
                                   rna_pc_num = 50)

subset_demo_seurat_1 <- Seurat::AddMetaData(
                                object = subset_demo_seurat_1,
                                metadata = pseudotime(subset_demo_cds_1),
                                col.name = "m3_pseudotime"
                                )

```


## Results: Monocle3 trajectory

```{r}
# Display results
# Check monocle3 trajectory
p_m3_pseudo_all <- plot_cells(subset_demo_cds_1,
                    color_cells_by = "pseudotime",
                    label_cell_groups=FALSE,
                    label_leaves=FALSE,
                    label_branch_points=FALSE,
                    graph_label_size=1.5)

wrap_plots(p_m3_pseudo_all)

func_save_images(image.object = list(wrap_plots(p_m3_pseudo_all)),
                 image.name = "mouse demo 1 - Subset - monocle3 trajectory",
                 image.path = saveTo,
                 h = 5, 
                 w = 6, 
                 r = 300, 
                 isHeatmap = F
                 )


# Check pseudotime
p_m3_celltype <- ggplot(subset_demo_seurat_1@meta.data,
                        aes(x = m3_pseudotime,
                        y = cell_type, colour = cell_type)) +
                        geom_point() +
                        geom_jitter(width = 0.1, height = 0.2) +
                        theme_gray() +
                        theme(legend.position = "none")

p_m3_celltype

func_save_images(image.object = list(p_m3_celltype),
                 image.name = "mouse demo 1 - Subset - monocle3 pseudotime",
                 image.path = saveTo,
                 h = 5,
                 w = 6,
                 r = 300,
                 isHeatmap = F
                 )

```



## Trajectory analysis - Slingshot
```{r}
# Slingshot pseudotime analysis

# Define function
func_slingshot <- function(input_seurat)
     {
  
      # group cells at early states as start_clus
  
      # This is just an example. Please define root cells that are biologically 
      # correct for your experiment.
      # Note: if ids is an empty object, then this function may generate error.
      # You can use this line instead:       

    	      #    demo_sce <- slingshot::slingshot(demo_sce, 
            #                                     clusterLabels = "cell_type",
            #                                     reducedDim = "PCA", 
            #                                     allow.breaks = F)

      # and remove the following four lines, "Seurat::Idents(input_seurat) <- ..",
      # "ids <- ..."
      # "input_seurat$cell_type <- .." and 
      # "input_seurat$cell_type[input_seurat$cell_type %in% ids] <- .."

      # for more information, please visit 
      # http://www.bioconductor.org/packages/release/bioc/vignettes/slingshot/inst/doc/vignette.html
  
      Seurat::Idents(input_seurat) <- "cell_type"

      ids <- unique(input_seurat$cell_type) %>% 
             .[grep(paste("CMP", 
                          "GMP", 
                          "HSC", 
                          "CD34", 
                          "stem", 
                          "iPS", 
                          "Myelocyte",
                          sep = "|"), 
               ., 
               ignore.case = T)]
      
      # group root cells
      input_seurat$cell_type[input_seurat$cell_type %in% ids] <- "start_clus"
      
      # convert Seurat to SingleCellExperiment class
      demo_sce <- Seurat::as.SingleCellExperiment(input_seurat)
         
      # Perform slingshot analysis
      temp <- tryCatch(
        {
          
         # PCA, tSNE and UMAP are all accepted for slingshot analysis. 
         # However, PCA has more dimensions. Hence, it's recommended by Slingshot Author.
         
         # Subset first 6 PC components
         reducedDim(demo_sce, 
                    type = "PCA", 
                    WithDimnames = TRUE) <- reducedDim(demo_sce,
                                                       type = "PCA")[, 1:6]
          
         demo_sce <- slingshot::slingshot(demo_sce, 
                                clusterLabels = "cell_type",
                                start.clus = "start_clus",
                                reducedDim = "PCA", 
                                allow.breaks = F)
        },
        error = function(e){
          
         # if there is an issue with PC selection, switch reduction method to UMAP
         demo_sce <- slingshot::slingshot(demo_sce, 
                                clusterLabels = "cell_type",
                                start.clus = "start_clus",
                                reducedDim = "UMAP", 
                                allow.breaks = F)
          
        }
      )
      
      # return SCE object
      return(temp)

} # end of function

# use function to get results
subset_demo_slingshot_1 <- func_slingshot(subset_demo_seurat_1)

pt_lineages <- slingshot::slingPseudotime(subset_demo_slingshot_1)

# add Slingshot results to the input Seurat object
lineages <- sapply(slingLineages(colData(subset_demo_slingshot_1)$slingshot), 
                   paste, 
                   collapse = " -> ")

subset_demo_seurat_1@meta.data[lineages] <- pt_lineages

```



## Results: Slingshot
```{r}
# display results

# display every lineage pseudotime
name_lineage <- colnames(subset_demo_seurat_1@meta.data)[grepl("start_clus",
                                                               colnames(subset_demo_seurat_1@meta.data))]

p_ss_1 <- list()

for (i in name_lineage) {
  
  p_ss_1[[i]] <- Seurat::FeaturePlot(subset_demo_seurat_1,  
                                     features = i) & 
                                     theme(legend.position="top") &
                                     scale_color_viridis_c() 
}

wrap_plots(p_ss_1, 
           ncol = 2)

func_save_images(image.object = list(wrap_plots(p_ss_1, 
                                                ncol = 2)),
                 image.name = "Subset - slingshot pseudotime (in Seurat)",
                 image.path = saveTo,
                 h = 15, 
                 w = 12, 
                 r = 100, 
                 isHeatmap = F
                 )


```


```{r}
# Display lineage curves
color <-  colorRamps::blue2green2red(length(unique(subset_demo_seurat_1$cell_type)))

names(color) <- unique(subset_demo_seurat_1$cell_type)

tiff(filename = paste(saveTo, 
                      "Subset - slingshot lineages.tiff", 
                      sep = "/"), 
     compression = "lzw", 
     width = 6, 
     height = 5, 
     units = "in", 
     res = 300)

{
par(mar = c(6,4,6,12) + 0.1, 
    xpd = T)
  
plot(reducedDim(subset_demo_slingshot_1, "PCA"), 
     pch=16, 
     cex = 0.5, 
     col = color[subset_demo_seurat_1$cell_type])

     title("Slingshot Lineages")

     legend("right",
            legend = names(color), 
            fill = color, 
            inset = c(-0.7,0))

     lines(SlingshotDataSet(colData(subset_demo_slingshot_1)$slingshot), 
           lwd = 2, 
           col = 'black', 
           type = 'lineages')
}

dev.off()

```


```{r}
# Check pseudotime for different cell types

p_ss_celltype <- list()

for (i in name_lineage) {
  
p_ss_celltype[[i]] <- ggplot(subset_demo_seurat_1@meta.data, 
                        aes(x = .data[[i]], 
                        y = cell_type, 
                        colour = cell_type)) +
                        geom_point() +
                        geom_jitter(width = 0.1, 
                                    height = 0.2) +
                        theme_gray() +
                        theme(legend.position = "none", 
                              plot.title = element_text(size = 1))
}

wrap_plots(p_ss_celltype, 
           ncol = 2)

func_save_images(image.object = list(wrap_plots(p_ss_celltype, 
                                                ncol = 1)),
                 image.name = "Subset - slingshot pseudotime",
                 image.path = saveTo,
                 h = 14, 
                 w = 5, 
                 r = 300, 
                 isHeatmap = F
                 )

```


## Cell-cell communication - CellChat
```{r}
# check existence of required libraries
if (!require(CellChat)) devtools::install_github("sqjin/CellChat")

# load library
library(CellChat)

# CellChat cell-cell communication analysis

# create function
func_cellchat <- function(input_seurat,
                          output_name)
     {
  
      # use cell type to grouop cells
      cc.object <- CellChat::createCellChat(object = input_seurat, 
                                            group.by = "cell_type")
 
      cc.object@DB <- CellChat::CellChatDB.mouse

      cc.object <- CellChat::subsetData(cc.object)
      
      cc.object <- CellChat::identifyOverExpressedGenes(cc.object)

      cc.object <- CellChat::identifyOverExpressedInteractions(cc.object)

      # cc.object <- CellChat::projectData(cc.object, PPI.mouse)

      cc.object <- CellChat::computeCommunProb(cc.object, 
                                               population.size = T, 
                                               raw.use = T)
 
      cc.object <- CellChat::computeCommunProbPathway(cc.object)

      cc.object <- CellChat::aggregateNet(cc.object)
      
      output <- CellChat::subsetCommunication(cc.object, 
                                              slot.name = "net")
      
      # Save the curated data to local drive
      temp_folder <- paste(saveTo, 
                           "CellChat output", 
                           sep = "/")

      ifelse(!dir.exists(file.path(temp_folder)), 
             dir.create(file.path(temp_folder)), FALSE)

      write.csv(x = output, 
                file = paste(temp_folder, 
                             paste0(output_name,".csv"), 
                             sep = "/"))

} # end of function

# use function to export files
# parameter "output_name" for function func_cellchat can any string, e.g. output_name <- "PBMCs_1".
# Here, Seurat project names are used.
func_cellchat(input_seurat = subset_demo_seurat_1, 
              output_name = Project(subset_demo_seurat_1))

```



## Save subsetted Seurat objects
```{r}
# Save Seurat objects to local drive
save(subset_demo_seurat_1,
     subset_demo1_DGEs,
     subset_demo_cds_1,
     subset_demo_slingshot_1,
     file = paste(saveTo, 
                  "subsetted demo Seurat objects (advanced analysis).RData", 
                  sep = "/"))

```


# Visualization

# cerebroApp
```{r}

# load libraries
library(cerebroApp)
library(InterCellar)
library(Seurat)
library(tidyverse)

```


```{r}
# Prepare file for CerebroApp export
# load cell cycle genes for mouse
# more details on how to get these genes can be found on this Seurat issue page on Github
# https://github.com/satijalab/seurat/issues/2493

load("BD-Demo-mouse-wta/cell_cycle_genes_mouse.RData")

# Build function
func_export_cerebro <- function(input_seurat,
                                DGE,
                                species,
                                enableEnrichedPathways,
                                saveTo = saveCerebroFolder,
                                output_name)
     {
  
      #  1) the maker genes can be tucked in here. Note: cerebroApp itself also has a
      #     function to calculate the differential gene expression, but it uses the 
      #     same method as the FindAllMarkers() in the Seurat package, which takes a
      #     relatively longer time.
      colnames(DGE$DGEs_cluster)[which(colnames(DGE$DGEs_cluster) == "cluster")] <- "cell_cluster"
      colnames(DGE$DGEs_cell)[which(colnames(DGE$DGEs_cell) == "cluster")] <- "cell_type"
  
      input_seurat@misc[["marker_genes"]][["cerebro_seurat"]][["cell_cluster"]] <- DGE$DGEs_cluster
      input_seurat@misc[["marker_genes"]][["cerebro_seurat"]][["cell_type"]] <- DGE$DGEs_cell
  
      # 2) build cluster tree
      # based on cell type annotation
      Seurat::Idents(input_seurat) <- "cell_type"

      input_seurat <- BuildClusterTree(
                           input_seurat,
                           dims = 1:10,
                           reorder = FALSE,
                           reorder.numeric = FALSE)

      input_seurat@misc$trees[["cell_type"]] <- input_seurat@tools$BuildClusterTree
      
      # based on Seurat clusters
      Seurat::Idents(input_seurat) <- "seurat_clusters"

      input_seurat <- BuildClusterTree(
                           input_seurat,
                           dims = 1:10,
                           reorder = FALSE,
                           reorder.numeric = FALSE)

      input_seurat@misc$trees[["seurat_clusters"]] <- input_seurat@tools$BuildClusterTree

      # 3) cell cycle analysis
      input_seurat <- CellCycleScoring(
                           input_seurat,
                           g2m.features = mm_g2m,
                           s.features = mm_s)

      input_seurat@misc$gene_lists$G2M_phase_genes <- mm_g2m

      input_seurat@misc$gene_lists$S_phase_genes <- mm_s

      # 4) calculate most expressed genes for each cell group
      input_seurat <- cerebroApp::getMostExpressedGenes(
                                  input_seurat,
                                  assay = 'RNA',
                                  groups = c('cell_type',
                                             'seurat_clusters')) # can add more

      # more examples can be found at
      #  https://romanhaa.github.io/cerebroApp/articles/cerebroApp_workflow_Seurat.html

      # 5) calculate MT and Rib genes with cerebroApp
      input_seurat <- addPercentMtRibo(
                               input_seurat,
                               organism = species, 
                               gene_nomenclature = 'name'
                  )
      
      
      # 6) calculate pathway enrichment with cerebroApp
      if (enableEnrichedPathways) {
        
          input_seurat <- getEnrichedPathways(
                            input_seurat,
                            marker_genes_input = 'cerebro_seurat',
                            adj_p_cutoff = 0.01,
                            max_terms = 100)
          }
      
      
      # 7) export the Seurat object as .crb format
      # Save the curated data to local drive
      ifelse(!dir.exists(file.path(saveTo)), 
             dir.create(file.path(saveTo)), 
             FALSE)

      cerebroApp::exportFromSeurat(
                  object =  input_seurat,
                  assay = "RNA",
                  slot = 'data',
                  file = paste(saveTo, 
                               paste0(output_name,
                                      "_cerebro_rna.crb"), 
                               sep = "/"),
                  experiment_name = "demo",
                  organism = species, 
                  groups = c("cell_type",
                             "seurat_clusters"),
                  nUMI = "nCount_RNA", 
                  cell_cycle = "Phase",
                  nGene = "nFeature_RNA",
                  add_all_meta_data = TRUE,
                  verbose = FALSE)

      if ("AB" %in% Seurat::Assays(input_seurat))
             {
              cerebroApp::exportFromSeurat(
                          object =  input_seurat,
                          assay = "AB", 
                          slot = 'data',
                          file = paste(saveTo, 
                                       paste0(output_name,
                                      "_cerebro_protein.crb"), 
                                       sep = "/"),
                          experiment_name = "demo",
                          organism = species,
                          groups = c("cell_type", 
                                     "seurat_clusters"),
                          nUMI = "nCount_AB", 
                          cell_cycle = "Phase",
                          nGene = "nFeature_AB", 
                          add_all_meta_data = TRUE,
                          verbose = FALSE)
             } # end of if statement

      # return Seurat object
      return(input_seurat)

} # end of function

# Use function to export file

# parameter "output_name" for function func_export_cerebro can be any string, 
# e.g. output_name <- "PBMCs_1".
# Here, Seurat project names are used.
subset_demo_seurat_1 <- func_export_cerebro(input_seurat = subset_demo_seurat_1, 
                        DGE = subset_demo1_DGEs, 
                        saveTo = saveTo,
                        species = "mm", 
                        enableEnrichedPathways = TRUE, # or FALSE to disable
                        output_name = Project(subset_demo_seurat_1))

```


## Save Seurat objects with cerebroApp content to local drive
```{r}
save(subset_demo_seurat_1,
     file = paste(saveTo,
                  "Seurat objects (cerebroApp).RData", 
                  sep = "/"))

```


## Data visualization with cerebroApp
```{r}
# launch cerebroApp

cerebroApp::launchCerebroV1.3()

```

## Data visualization with InterCellar (cell-cell communication)

```{r}
# activate interactive user interface.
InterCellar::run_app()

# more information can be found at https://github.com/martaint/InterCellar/ 

```





