---
title: "Packages installation"
author: "Wenyan Li"
date: "2022-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Computer for script testing

Computer configuration:
Processor	Intel(R) Xeon(R) Bronze 3204 CPU @ 1.90GHz   1.90 GHz  (2 processors)
Installed RAM	32.0 GB (31.6 GB usable)
System type	64-bit operating system, x64-based processor

Edition	Windows 10 Pro for Workstations
Version	21H2
Installed on	18/03/2021
OS build	19044.1889


## Software needed for the script

Download software
RStudio version used: RStudio-2022.07.2-576.exe
(Link to download: https://www.rstudio.com/products/rstudio/download/)

R version used: R-4.2.1-win.exe
(Link to download: https://cran.r-project.org/bin/windows/base/old/4.2.1)

Rtools version used: rtools42-5355-5357.exe
(Link to download: https://cran.r-project.org/bin/windows/Rtools/rtools42/rtools.html)

Note: if the operation system is MacOS, download the software version for MacOS


## Install RStudio, R and Rtools
Install them with default settings



## Open RStudio and install packages

1. Install tools for package installation
```{r}
install.packages('BiocManager')
install.packages('remotes')
install.packages('devtools')
```

At the beginning of installing devtools, a window may pop up and ask “Do you want to install from sources the package which needs compliation?”. Click “Yes” to continue. Note: The installation of this package may take a longer time than the other packages.

2. Install packages from CRAN
```{r}
install.packages('magick')
install.packages('tidyverse')
install.packages('ggplot2')
install.packages('Seurat')
install.packages('harmony')
install.packages('patchwork')
install.packages('R.utils')
install.packages('terra')
install.packages('sf')
install.packages('rstudioapi')
install.packages('colorRamps')
install.packages('viridis')
install.packages('pheatmap')
install.packages('ape')
```

During installation of the 'Seurat' package, a window may pop up and ask “Do you want to install from sources the package which needs compliation?”. Click “Yes” to continue.


3. Install packages from BioConductor
```{r}
BiocManager::install('EBImage')
BiocManager::install('celldex')
BiocManager::install('SingleR')
BiocManager::install('Slingshot')
BiocManager::install('InterCellar')
BiocManager::install('romanhaa/cerebroApp')
BiocManager::install('flowCore')
BiocManager::install("ComplexHeatmap")
```

4. Install packages from GitHub
```{r}
devtools::install_github('chris-mcginnis-ucsf/DoubletFinder')
devtools::install_github('satijalab/seurat-wrappers')
devtools::install_github('cole-trapnell-lab/monocle3')
devtools::install_github('sqjin/CellChat')
devtools::install_github('immunogenomics/presto')
```














